---
title: "Running Likelihood Ratio Test"
author: "Elena Venable"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
---
## 1. Use Likelihood to Generate Genotype Probabilities
The `likelihood` can be used to only calculate genotype probabilities if the gt is omitted as shown below.

```{r}
# generates likelihood probabilities given allele frequencies
library(SNPcontam)
af <- c(SNP1 = .2, SNP2 = .6, SNP3 = .25, SNP4 = 0.8)
likelihood(af)
```
The list componenets contam and clean are Null because no genotype was entered.

## 2. Use Function to Randomly Generate Genotypes
The `random_gene` can next be used to generate genotypes based on the genotype frequencies computed usking `likelihood`.

```{r}
# generates 5 random genotypes
Ran <- random_gene(5,af)
```

`rclean` includes genotypes generated assuming no contamination while `rcontam` includes those generated assuming contamination.

## 3. Use Function to Compute Lilkelihoods of Random Genotypes
The `likehood` function can be used to compute the likelihoods of genotypes.
```{r}
# likelihood for the non contaminated samples
L <- likelihood(af,Ran$rclean)
L
# likelihood for the contaminated samples
LC <- likelihood(af, Ran$rcontam)
```
LC will have a similar output to L, but for the contaminated assumption.

## 4. Use Function to Compute Likelihood Ration
The function `lratio` can be used to compute the likelihood ratio for each individual.
```{r}
# ratios for the non contaminated individuals
ratio <- lratio(L$clean, L$contam,af)
ratio

# ratios for the contaminated individuals
ratio_c <- lratio(LC$clean, LC$contam,af)
ratio_c
```

## 5. Use Function to Compute Heterozygousity
The function `hetero` can be used to compute the proportion of heterozygous loci in an individual.
```{r}
# proportion for non contaminated individuals
hetero <- hetero(Ran$rclean,af)
hetero

# proportion of contamintated individuals
hetero_c <- hetero(Ran$rcontam, af)
hetero_c
```

## 6. Use Function to Plot ROC Curves
```{r}
#subplots
par(mfrow=c(1,2))
# plots the ROC for the ratio test
ROC(10,ratio[[1]], ratio_c[[1]])

#plots the ROC for the heterozygosity test
ROC(10,hetero[[1]],hetero_c[[1]])
```



Histograms of the different likelihood ratio and heterozygosity distributions can also be plotted:
```{r}
par(mfrow=c(1,2))
# likelihood ratio test
hist(ratio[[1]],col=rgb(1,0,0,0.5),xlim = c(-5,10),main="Likelihood Ratio Histogram")
hist(ratio_c[[1]],col=rgb(0,0,1,0.5),add=T)

# heterozygousity proportions
hist(hetero[[1]],col=rgb(1,0,0,0.5),xlim = c(0,1),main="Heterozygousity Proportion Histogram")
hist(hetero_c[[1]],col=rgb(0,0,1,0.5),add=T)
```

## Example of Output
Below is code that synthesizes the steps above.  There are 20 randomly generated alle frequencies and 1000 randomly generated individuals.
```{r}
af <- runif(20,0,1)
Ran <- random_gene(1000,af)
L <- likelihood(af,Ran$rclean)
LC <- likelihood(af, Ran$rcontam)
ratio <- lratio(L$clean, L$contam,af)
ratio_c <- lratio(LC$clean, LC$contam,af)
hetero <- hetero(Ran$rclean,af)
hetero_c <- hetero(Ran$rcontam, af)
par(mfrow=c(2,2))
ROC(100,ratio[[1]], ratio_c[[1]])
hist(ratio[[1]],col=rgb(1,0,0,0.5),xlim = c(-40,20),main="Likelihood Ratio Histogram")
hist(ratio_c[[1]],col=rgb(0,0,1,0.5),add=T)
ROC(100,hetero[[1]],hetero_c[[1]])
hist(hetero[[1]],col=rgb(1,0,0,0.5),xlim = c(0,1),main="Heterozygousity Histogram")
hist(hetero_c[[1]],col=rgb(0,0,1,0.5),add=T)
```